PhD Title \\ PhD Thesis

### 2024-06-25 Latex_files/Thesis/chapters/chapter 1/main.tex
updates to ch1 and ch3

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain miscible while 
below a certain threshold concentration ($C_{eq}$) 
- the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate 
the two phases. Once $C_{eq}$ is exceeded the there is
a potential difference driving the solution to separate the 
two phases. Since different combinations of solute and solvent
will have different threshold concentrations, researchers often 
instead measure the ratio between the solute and solvent by using 
'supersaturation', one form of measuring supersaturation is shown 
below \cite{Mullin2001}:
\begin{align}
	S(T) = \frac{C_{solute}}{C_{eq}(T)}
\end{align}
Where $C_{solute}$ is just the raw solute concentration, $C_{eq}(T)$
is the threshold concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. In particular 
there is a growing desire for a unified theory to 
describe the multiple paths that nucleation can take.

\subsection{Primary vs Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where consistent 
residence times are desirable.

In contrast, secondary nucleation is the result of a initial 
seed crystal inducing further nucleation within the bulk solution
\cite{Botsaris1976}, in absence of a seed crystal nucleation is
near impossible. The research of Secondary Nucleation is fascinating
for its own merits, not only is it speculated that the supersaturation 
barrier is non-existent \cite{Cashmore2022}

@@

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}
Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
\end{figure}

@@

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, as in situ techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation.  It can be understood somewhat by Oswald's rule, where he found that for supersaturated solutions with multiple potential polymorphs, the general tendency was for the least stable polymorph to form first, and then over time the more stable polymorph to form afterwards. This was explained as the system preferring to take a quick step to an intermediate state of lower free energy first, before eventually tending towards the lowest possible free energy state. Likewise for two-step nucleation, while nucleation is by far the pathway with lowest free energy, it is still possible for the solutes to find an intermediate state such as a cluster of solute molecules prior to nucleation. The idea of only two steps is merely a convention and so some researchers prefer to refer to it as multi-step nucleation, where any number of intermediate structures can occur between the base solution and final nucleation event.

>>>>

\section{Nucleation}
Nucleation is an example of a binary phase separation, 
where a dilute phase is miscible in a bulk phase, more 
often called the solute and solvent respectively. Because 
of thermodynamics, the two can only remain in equilibrium while 
below a specific concentration ($C_{eq}$) - below which 
the chemical potential $\mu$ for a miscible solution 
is greater than the potential required to separate 
the two phases. Once $C_{eq}$ is exceeded there is a chemical
 potential difference driving the solution to separate the 
two phases. Since different combinations of solute and solvent
will have different equilibrium concentrations, researchers often 
instead measure the ratio between the solute and solvent by using 
'supersaturation', one form of measuring supersaturation is shown 
below \cite{Mullin2001}:
\begin{align}
	\label{eq:supersaturation}
	S(T) = \frac{C_{sol}}{C_{eq}(T)}
\end{align}
Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
is the equilibrium concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation
of the system.

\subsection{Primary vs Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where short and consistent 
residence times are desirable. At a small scale one can estimate the 
nucleation rate by making repeated measurements of sample solutions 
and seeing how many have nucleated after a given time, giving us a 
Poisson distribution.

\begin{align}
	P(t) = 1 - exp\left[-JV(t-t_g)\right] = \frac{M^*(t)}{M}
\end{align}

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$
is the volume of the individual samples, and $M^*(t)$ \& $M$ are the 
number of nucleated samples and the total number of samples used respectively.
While this is useful for studying the effects of different parameters
at a small scale, for industrial applications there are so many
additional factors at play that can throw off this estimation.

In contrast, secondary nucleation is the result of a initial 
seed crystal inducing further nucleation within the bulk solution
\cite{Botsaris1976}, in absence of a seed crystal nucleation is
near impossible. The research of Secondary Nucleation is fascinating
for its own merits; is it speculated that the supersaturation 
barrier is non-existent \cite{Cashmore2022}, and furthermore there 
are a multitude of mechanisms that can induce secondary 
nucleation making it imperative that researchers can control the
nucleation rate at an industrial scale.

\begin{figure}[h!]
	\includegraphics[width=\linewidth]{secondary_nucleation.jpg}
	\caption{Secondary Nucleation mechanisms, classified by Agrawal and Paterson \cite{Agrawal2015}}
\end{figure}

@@

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}
Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\label{fig:free_energy}
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
\end{figure}

@@

\subsection{Two Step Nucleation}
The two step nucleation theory is an extension to the CNT, colloid simulations showed that short range (such as in proteins \cite{Wolde1997, Gliko2005}) interactions allow for the formation of a liquid-liquid metastable phase from which a new solid phase could form \cite{Anderson2002, Karthika2016}. \textit{in situ} techniques for studying nucleation several papers reported the presence of stable liquid-like clusters that formed prior to nucleation \cite{Savage2009, Wolde1997, Soga1999}.  It can be understood by Oswald's rule \cite{Ostwald1897}, which says that any crystallising system does not 
take the path to the lowest possible energy state but instead first transitions to the state with the smallest free energy barrier, further transitions can still occur but this method minimises the overall free energy cost. If we consider the free energy diagram from before (fig.~\ref{fig:free_energy}) we can consider a case where the supersaturation is low ($S\approx1.01$) then a dense liquid droplet will have a lower chemical potential than the surrounding fluid

\section{Crystallisation methods}
\subsection{Cooling Crystallisation}
For some binary mixtures the supersaturation is heavily dependent on 
the solution temperature, therefore a simple method of producing crystals 
is by cooling the mixture to induce crystal formation. At ambient 
temperatures the solute concentration is too high to be fully incorporated
into the solution ($S\gg 1$), after heating however the solute is fully dissolved
into the solution ($S<1$). Now as the mixture is allowed to cool to room
temperature the supersaturation will increase until crystal formation begins,
the rate of cooling drastically influencing the size and number of crystals
produced.

If $dT/dt$ is high then the final product will consist of large 
crystal and be low in number, as the nucleation rate is directly related
to the supersaturation only a handful of nuclei can form before the remaining
solute grows onto the surface. If $dT/dt$ is low then the final product will 
consist of smaller crystals and be far more numerous, as the supersaturation 
is so large that the new nuclei are forming continuously. Between these two 
extremes, one can define the meta-stable zone width, a region in the 
temperature-concentration phase space where both nucleation and crystal growth
can be reliably controlled. The lower limit being given by the solubility curve
of the binary mixture, and the upper limit being defined by the metastable zone.

\begin{figure}[h!]
	\label{fig:MSZW}
	\includegraphics[width=\linewidth]{MSZW.png}
	\caption{Typical Concentration vs Temperature plot with the solubility curve (black) and metastable zone (red). 3 different cooling curves are shown as well: a high rate of cooling (orange) shows the mixture quickly dropping below the solubility curve resulting in no new nuclei forming; a low rate of cooling (green) shows the mixture exceeding the metastable zone, where now nucleation occurs freely; and a typical cooling rate (blue) shows how a typical cooling crystalliser will operate, sitting in the middle of the other two curves.}
\end{figure}

The viability of cooling crystallisation is dependent on the meta-stable 
zone width, too narrow and the process is difficult to control, too wide 
and the crystal growth rate may be insufficient for the desired outcome.

### 2024-06-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch1 and ch3

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution. The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results. It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}. Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}. These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}.

\section{Optical Tweezer Equipment}
In general, all optical tweezers require a low power laser driver, two microscope objectives (one for trapping and one for imaging), a means of controlling the position of the loaded sample, and a means of isolating vibration signals. The laser used for this project was a 1064 nm near infrared laser - provided by CNI Lasers – that has an adjustable power supply to vary the energy output of the laser. Experimental work has shown that the trapping efficiency increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ where $D_{obj}$ is the diameter of the objective aperture. To expand the beam front we utilise a Galilean beam expansion arrangement (indicated by $L_1$, and $L_2$) as recommended for high power laser applications. In our initial experiments the beam expansion provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then the 4f correlator provides a further $1.25\times$ magnification. Afterwards the laser is passed through a dichroic mirror that reflects oncoming infrared light but allows the passage of LED light through, this is to prevent the laser from interfering with the CCD imaging. By increasing the numerical aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth is reduced due to spherical aberrations. While it is possible to increase the trapping depth by adjusting the objective's tube length this approach is incompatible with our trapping arrangement. The condenser objective refocuses the scattered laser light and also provide an aperture for an imaging LED to illuminate the focal plane. Samples are loaded onto a piezo driven table to that is inserted between the trapping and condensing objectives; the piezo drivers allow for sub-micron control of the sample position to a degree as small as a 10 nm. To detect and monitor the position of a trapped particle a quadrant photo diode (QPD) was utilised.

\subsection{Position detection methods}
There are several methods for monitoring the position of 
A QPD is a typical example of a position detection system commonly used 
in optical tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is directed onto the QPD, with the maximum 
intensity mapping to the particle's centre of mass. By summing the 
horizontal and vertical quadrants together the particle's centre of mass 
is tracked in the x-y plane. While the tracking is accurate the QPD 
needs to be calibrated in order to convert the voltage signals into 
actual distance reading.

>>>>

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the PhD was to investigate the possibility of using optical tweezing as a means of initiating and controlling nucleation by generating fluid flow within a small droplet of supersaturated solution. The goal of which would be to allow the understand the influence of shearing on nucleation at a micro level as compared to larger scale results. It has been shown that for macro-scale systems, the likelihood of nucleation increases to a maximum value under increased shearing \cite{Debuysschere2023, Mura2016}. Theoretical research into the matter identified two competing effects that effect a crystal in moving fluid fields; firstly, nucleation is enhanced due to the increased mass transfer of solute material; and secondly, shear flow against the crystal surface leading to a decrease in growth \cite{Mura2016}. These two competing effects are validated by experimental work using glycine solutions, showing that beyond a certain shear rate the nucleation rate is reduced \cite{Debuysschere2023}. In this chapter I outline the optical tweezer equipment used during this PhD, the initial attempts made to induce nucleation via optical rotation, and the impact of a moving beam on the crystal growth of a newly formed nucleus.

\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, two
microscope objectives (one for trapping and one for imaging),
and a means of controlling the position of the loaded sample. 
While there are other pieces of equipment used in modern tweezer
experiments these are the bare minimum requirements for any 
optical tweezer. The laser used for this project was a 1064 
nm near infrared laser - provided by CNI Lasers – that has 
an adjustable power supply to vary the energy output of the 
laser - the remaining optics where supplied by Thor Labs.
Experimental work has shown that the trapping efficiency 
increases with beam diameter up until it exceeds $\frac{2}{3}D_{obj}$ 
\cite{kim2003dependence} where $D_{obj}$ is the diameter
of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated 
by $L_1$, and $L_2$) as recommended for high power laser 
applications. In our initial experiments the beam expansion 
provides a $4\times$ magnification; however, in later experiments where we utilise a galvano-mirror the beam expansion is $3\times$ and then 
the 4f correlator provides a further $1.25\times$ magnification.
It should be noted that the galvano-mirror requires the use 
of a Keplerian beam expansion arrangement which reduces
the transmitted laser power due to localised heating of the air.
Afterwards the laser is passed through a dichroic mirror that 
separates incoming infrared and visible light, this is to 
prevent the laser from damaging the CCD camera used for imaging
the trapping plane. The laser is then focused to a diffraction
limited spot by a 1.25 NA objective. By increasing the numerical
aperture of the objective, the gradient force at the focal point is increased; the trade-off being that the for higher NA values the trapping depth 
is reduced due to spherical aberrations. While it is possible 
to increase the trapping depth \cite{Reihani2006} by adjusting 
the objective's tube length this approach is incompatible with 
our trapping arrangement. A 0.25 NA condenser objective refocuses
the scattered laser light and also provide an aperture for an 
imaging LED to illuminate the focal plane. Samples are loaded 
onto a piezo driven table to that is inserted between the trapping
and condensing objectives; the piezo drivers allow for sub-micron
control of the sample position to a degree as small as a 10 nm. 
To detect and monitor the position of a trapped particle a 
quadrant photo diode (QPD) was utilised.

\begin{figure}
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively.}
\end{figure}

\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, 
a position detection system must be utilised. As shown 
A QPD is a typical example of a position detection system commonly used 
in optical tweezers due to their high sampling rate, high degree of 
precision, and ease of set up. The QPD is constructed of four photo 
diodes assembled in a quadrant formation, when a particle is trapped the 
interference pattern produced is directed onto the QPD, with the maximum 
intensity mapping to the particle's centre of mass. By summing the 
horizontal and vertical quadrants together the particle's centre of mass 
is tracked in the x-y plane. While the tracking is accurate the QPD 
needs to be calibrated in order to convert the voltage signals into 
actual distance reading.
